Formalizing Stack Safety as a Security Property

### 2022-10-19 submission5/noexceptions.tex
Small changes
More example discussion in definitions

\definition A system enjoys well-bracketed control flow if, for all states
\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context)\),
\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).

@@

\definition Let \((\mach,\context)\) be a combined state with
\(\context = (V,\sigma)\), and
\(\components = \{\component | V ~ \component = \sealed\}\).
Let \(P\) be a predicate that holds on a state \((\mach',\context')\) if
\(\mach' ~ \parallel ~ \Delta(\mach,\mach') \cap \components\).
Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \uparrow P\) holds
on \((\mach,\context)\).

@@

@@

>>>>

\definition A system enjoys well-bracketed control flow if, for all states
\((\mach,(\_,(\_,\addr_{ret},\addr_{sp})) \stepstounder{\mathbf{return}} (\mach',\context,\obs)\),
\(\mach'[\PCname] = \addr_{ret}\) and \(\mach'[\SP] = \addr_{sp}\).

Consider again the attack in Figure \ref{subfig:WBCF}: the attacker adds
20 to the return address and then returns, thus bypassing the check and publishing
{\tt secret}. During the call step, we record in the context that the return to the
caller should have the program counter at the next instruction (address 20) and the
stack pointer restored to address -20. When the return occurs, we see that in the
state following it, the program counter is 40 -- a violation of WBCF.

To demonstrate why WBCF also cares about the stack pointer, consider an attack which
returns with the stack pointer at -12. Now, compare the layout of the stack relative
to the stack pointer as it should be (top) to the one that results by increasing the
stack pointer before return (bottom):

\begin{tabular}{| l | l | l | l | l | l | l |}
  \multicolumn{3}{r}{\(\SP \downarrow\)} \\
  \hline
  & & res & sens & secr & ra & ra \\
  \hline
  \hline
  res & sens & secr & ra & ra & & \\
  \hline
\end{tabular}

In this scenario, when attempting to read {\tt sensitive}, {\tt main} will
read part of the return address, and then it will attempt to publish
{\tt res}, but instead will publish {\tt secret}!

Alternatively, we can formalize well-bracketed control flow using \(\uparrow\).

\definition For a machine state \(\mach\), let
\(ret ~ (\mach,\context) = \{\mach' | \mach'[\SP] = \mach[\SP] \land \mach'[\PCname] = \mach[\PCname]\}\).
A system enjoys well-bracketed control flow if, for all states 
\((\mach,\context) \stepstounder{\overline{\psi}} (\mach',(V,\sigma),\obs)\) with
\(\mathbf{call} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\) or
\(\mathbf{tailcall} ~ \addr ~ \overline{\reg} ~ \overline{sa} \in \psi\),
\(|\sigma| \uparrow ret(\mach')\).

Once again we can think about our example attack. When the call occurs,
we define a predicate \(ret ~ (\mach,\context)\) on states that holds when the program counter
and stack pointer have the appropriate values. Then we use apply the \(\uparrow\)
operator to this predicate with the depth of the call stack {\it after the call} and
the machine state {\it before the call}.

After the call, the call stack is one higher than it was before the call,
and if the callee makes further calls, it will continue to be incremented.
Each return will similarly decrement it, until it is finally lower than the
initial state. This is the return that matches the original call, and so
the resulting state is when we require \(ret(\mach)\) to hold.

@@

\definition Let \((\mach,\context)\) be a combined state with
\(\context = (V,\sigma)\), and
\(\components = \{\component | V ~ \component = \sealed\}\).
Let \(P ~ (\mach',\context')\) be a predicate that holds on a state
\(\Delta(\mach,\mach')\) is irrelevant in \((\mach', \context')\).
Then \(\mach\) enjoys {\it return-time integrity} if \(|\sigma| \uparrow P\) holds
on \((\mach,\context)\).

@@

The example in Figure \ref{subfig:integrity} modifies the value of {\tt sensitive},
which is \(\sealed\). At the call, we define the predicate \(P\), which holds on states
where any \(\sealed\) elements that changed since the call are all irrelevant.
As {\tt sensitive} is such an element (conveniently, the only one), the property
has not yet been violated at the returned state \((\mach,\context)\). To determine whether this
attack has violated stack integrity, we must consider an arbitrary variant state
\(\nach = \mach[\addr_{\mathtt{sensitive}} \mapsto \word]\) for some \(\word\).
The property is violated if the traces of \((\mach,\context)\) and \((\nach,\context)\)
differ.

In our case, {\tt publish} produces an observation, and if \(\word \not = 42\),
that observation in the trace from \((\nach,\context)\) will be {\tt res}, which is 0.
On the other hand, \((\mach,\context)\) will produce an event from {\tt secret},
which may differ. So, integrity is indeed violated.

@@

Confidentiality is more complicated. We begin with a state immediately following
a call. Let that be \((\mach,\context)\). We suppose an arbitrary variant state,
\((\nach,\context)\), which may vary any element that is \(\sealed\) or \(\unsealed\).
That is, any element that is not used legitimately to pass arguments. Caller confidentiality
therefore can be thought of as the callee's insensitivity to parts of its initial states
that are not part of the caller-callee interface.

In the example in Figure \ref{subfig:direct}, this variation will quickly result
in differing events, as {\tt publish} is called with arguments that may have different
values. This is clearly a violation of confidentiality, in the form of internal
confidentiality.

When the variants return, however, the situation becomes more complicated.
Consider Figure \ref{subfig:indirect}. The value of {\tt secret} is been extracted
and placed in {\tt a0}, the return value register. Suppose that we have \(\mach\)
and \(\nach\) as above, and they step, respectively, to \(\mach'\) and \(\nach'\),
which each follow a return. We must identify which elements of the state
contain leaked confidential data.

An element contains confidential data, according to the definition of non-interference, 
if it differs between the states. But some of those data belong to the original
elements which are accessible to the caller and may now influence future execution
legitimately. We are looking specifically for those elements that have {\it changed}
from call to return, and that vary between the variants. As with integrity, we then
must determine whether those leaks actually translate into events. If the set of
such elements is not irrelevant, confidentiality is violated.
